---
title: "WCGNA Analysis of At, Nt and Vv"
---

### Install all needed packages for WCGNA to work

```{r}
install.packages("RSQLite")
BiocManager::install("GO.db")
BiocManager::install("WGCNA")
BiocManager::install("impute")

```

### Import libraries

```{r}
library(RSQLite)
library(WGCNA)
library(dplyr)
library(tidyverse)
library(magrittr)
library(ggplot2)
library(cluster)
```

### Load in normalised TPM data

```{r}
at_s1_tpm <- read.delim("at_s1_tpm.txt")
at_s2_tpm <- read.delim("at_s2_tpm.txt")
at_s3_tpm <- read.delim("at_s3_tpm.txt")

nt_s1_tpm <- read.delim("nt_s1_tpm.txt")
nt_s2_tpm <- read.delim("nt_s2_tpm.txt")
nt_s3_tpm <- read.delim("nt_s3_tpm.txt")


vv_p1_tpm <- read.delim("vv_p1_tpm.txt")
vv_p2_tpm <- read.delim("vv_p2_tpm.txt")

#remove gene id column
#rownames(at_s1_tpm) <- at_s1_tpm$GeneId
#at_s1_tpm <- at_s1_tpm[, -1]
#rownames(at_s2_tpm) <- at_s2_tpm$GeneId
#at_s2_tpm <- at_s2_tpm[, -1]
#rownames(at_s3_tpm) <- at_s3_tpm$GeneId
#at_s3_tpm <- at_s3_tpm[, -1]

rownames(nt_s1_tpm) <- nt_s1_tpm$GeneId
nt_s1_tpm <- nt_s1_tpm[, -1]
rownames(nt_s2_tpm) <- nt_s2_tpm$GeneId
nt_s2_tpm <- nt_s2_tpm[, -1]
rownames(nt_s3_tpm) <- nt_s3_tpm$GeneId
nt_s3_tpm <- nt_s3_tpm[, -1]

rownames(vv_p1_tpm) <- vv_p1_tpm$GeneId
vv_p1_tpm <- vv_p1_tpm[, -1]
rownames(vv_p2_tpm) <- vv_p2_tpm$GeneId
vv_p2_tpm <- vv_p2_tpm[, -1]
```

### Merge data sets for each species

```{r}
nt_tpm <- cbind(nt_s1_tpm, nt_s2_tpm, nt_s3_tpm)


at_tpm <- merge(at_s1_tpm, at_s2_tpm, by = 'GeneId')
at_tpm <- merge(at_tpm, at_s3_tpm, by = 'GeneId')
rownames(at_tpm) <- at_tpm$GeneId
at_tpm <- at_tpm[, -1]

vv_tpm <- cbind(vv_p1_tpm, vv_p2_tpm)

colnames(nt_tpm) <- c('S1_M1', 'S1_M2', 'S1_M3', 'S1_M4', 'S1_M5','S1_X1', 'S1_X2', 'S1_X3', 'S1_X4', 'S1_X5','S2_M1', 'S2_M2', 'S2_M3', 'S2_M4', 'S2_M5','S2_X1', 'S2_X2', 'S2_X3', 'S2_X4','S2_X5','S3_M1', 'S3_M2', 'S3_M3', 'S3_M4', 'S3_M5','S3_X1', 'S3_X2', 'S3_X3', 'S3_X4', 'S3_X5')

colnames(at_tpm) <- c('S1_M1', 'S1_M2', 'S1_M3', 'S1_M4', 'S1_M5','S1_X1', 'S1_X2', 'S1_X3', 'S1_X4', 'S1_X5','S2_M1', 'S2_M2', 'S2_M3', 'S2_M4', 'S2_M5','S2_X1', 'S2_X2', 'S2_X3', 'S2_X4','S2_X5','S3_M1', 'S3_M2', 'S3_M3', 'S3_M4', 'S3_M5','S3_X1', 'S3_X2', 'S3_X3', 'S3_X4', 'S3_X5')

colnames(vv_tpm) <- c('S2_M1', 'S2_M2', 'S2_M3','S2_X1', 'S2_X2', 'S2_X3','S3_M1', 'S3_M2', 'S3_M3', 'S3_X1', 'S3_X2', 'S3_X3')
```

### Remove 0 counts so that we can get higher average reads

```{r}
remove_zero <- function(df){
  zero_matrix <- df
  all_zero_rows <- apply(zero_matrix, 1, function(row) all(row == 0))
  zero_ids <-  df[all_zero_rows, 1]
  df_filt <- df[!df[, 1] %in% zero_ids, ]
  
  return(df_filt)
}

at_tpm_filt <- remove_zero(at_tpm) 
nt_tpm_filt <- remove_zero(nt_tpm)
vv_tpm_filt <- remove_zero(vv_tpm)
```

### \*Log transform TPM values so that they are normally distributed

```{r}
#Log Transform
at_log <- log2(at_tpm_filt + 1)

nt_log <- log2(nt_tpm_filt + 1)

vv_log <- log2(vv_tpm_filt + 1)
```

### Filter down Nt and Vv to fit the number of At genes

```{r}
core_orth <- read.delim('core_orth.txt')

core_orth$at <- gsub("\\.[0-9]+$", "", core_orth$at)

#Filter down on each list to find the common remaining non-zero read genes
core_orth <- core_orth[core_orth$nt %in% rownames(nt_tpm_filt), ]
at_tpm_filt <- at_tpm_filt[rownames(at_tpm_filt) %in% core_orth$at, ]
vv_tpm_filt <- vv_tpm_filt[rownames(vv_tpm_filt) %in% core_orth$vv, ]

#Now filter based on at
core_orth <- core_orth[core_orth$at %in% rownames(at_tpm_filt), ]
nt_tpm_filt <- nt_tpm_filt[rownames(nt_tpm_filt) %in% core_orth$nt, ]
vv_tpm_filt <- vv_tpm_filt[rownames(vv_tpm_filt) %in% core_orth$vv, ]

#Now filter based on vv
core_orth <- core_orth[core_orth$vv %in% rownames(vv_tpm_filt), ]
nt_tpm_filt <- nt_tpm_filt[rownames(nt_tpm_filt) %in% core_orth$nt, ]
at_tpm_filt <- at_tpm_filt[rownames(at_tpm_filt) %in% core_orth$at, ]

write.table(core_orth, "filtered_core_orth.tsv", sep = "\t", quote = FALSE)
```

### Combine the data sets

```{r}
#replace the gene names with the at genes so we can keep the same columns together when combining them
vv_tpm_filt_names <- cbind(vv_tpm_filt, rownames(vv_tpm_filt))
vv_tpm_filt_names <- vv_tpm_filt_names %>%
  rename('vv' = names(vv_tpm_filt_names)[ncol(vv_tpm_filt_names)])
                     
vv_tpm_with_pairs <- vv_tpm_filt_names %>%
  left_join(core_orth, by = "vv")
rownames(vv_tpm_with_pairs) <- vv_tpm_with_pairs$at
vv_tpm_with_pairs <- vv_tpm_with_pairs[, -c(13, 14, 15)]

#Nt
nt_tpm_filt_names <- cbind(nt_tpm_filt, rownames(nt_tpm_filt))
nt_tpm_filt_names <- nt_tpm_filt_names %>%
  rename('nt' = names(nt_tpm_filt_names)[ncol(nt_tpm_filt_names)])
                     
nt_tpm_with_pairs <- nt_tpm_filt_names %>%
  left_join(core_orth, by = "nt")
rownames(nt_tpm_with_pairs) <- nt_tpm_with_pairs$at
nt_tpm_with_pairs <- nt_tpm_with_pairs[, -c(31, 32, 33)]


combined_df <- merge(at_tpm_filt, nt_tpm_with_pairs, by = 'row.names')
rownames(combined_df) <- combined_df$Row.names
combined_df$Row.names <- NULL

combined_df <- merge(combined_df, vv_tpm_with_pairs, by = 'row.names')
rownames(combined_df) <- combined_df$Row.names
combined_df$Row.names <- NULL

names(combined_df) <- c('A_S1_M1', 'A_S1_M2', 'A_S1_M3', 'A_S1_M4', 'A_S1_M5','A_S1_X1', 'A_S1_X2', 'A_S1_X3', 'A_S1_X4', 'A_S1_X5','A_S2_M1', 'A_S2_M2', 'A_S2_M3', 'A_S2_M4', 'A_S2_M5','A_S2_X1', 'A_S2_X2', 'A_S2_X3', 'A_S2_X4','A_S2_X5','A_S3_M1', 'A_S3_M2', 'A_S3_M3', 'A_S3_M4', 'A_S3_M5','A_S3_X1', 'A_S3_X2', 'A_S3_X3', 'A_S3_X4', 'A_S3_X5', 'N_S1_M1', 'N_S1_M2', 'N_S1_M3', 'N_S1_M4', 'N_S1_M5','N_S1_X1', 'N_S1_X2', 'N_S1_X3', 'N_S1_X4', 'N_S1_X5','N_S2_M1', 'N_S2_M2', 'N_S2_M3', 'N_S2_M4', 'N_S2_M5','N_S2_X1', 'N_S2_X2', 'N_S2_X3', 'N_S2_X4','N_S2_X5','N_S3_M1', 'N_S3_M2', 'N_S3_M3', 'N_S3_M4', 'N_S3_M5','N_S3_X1', 'N_S3_X2', 'N_S3_X3', 'N_S3_X4', 'N_S3_X5', 'V_S2_M1', 'V_S2_M2', 'V_S2_M3','V_S2_X1', 'V_S2_X2', 'V_S2_X3','V_S3_M1', 'V_S3_M2', 'V_S3_M3', 'V_S3_X1', 'V_S3_X2', 'V_S3_X3')
```

### Averaged Replicates

```{r}
numeric_combined_df <- combined_df

# Create an empty dataframe to store averages
averaged_df <- data.frame(matrix(ncol = 16, nrow = 0))

for(row in 1:nrow(numeric_combined_df)){
  a_ms1 <- mean(unlist(unname(numeric_combined_df[row, 1:5])))
  a_xs1 <- mean(unlist(unname(numeric_combined_df[row, 6:10])))
  a_ms2 <- mean(unlist(unname(numeric_combined_df[row, 11:15])))
  a_xs2 <- mean(unlist(unname(numeric_combined_df[row, 16:20])))
  a_ms3 <- mean(unlist(unname(numeric_combined_df[row, 21:25])))
  a_xs3 <- mean(unlist(unname(numeric_combined_df[row, 26:30])))
  
  n_ms1 <- mean(unlist(unname(numeric_combined_df[row, 31:35])))
  n_xs1 <- mean(unlist(unname(numeric_combined_df[row, 36:40])))
  n_ms2 <- mean(unlist(unname(numeric_combined_df[row, 41:45])))
  n_xs2 <- mean(unlist(unname(numeric_combined_df[row, 46:50])))
  n_ms3 <- mean(unlist(unname(numeric_combined_df[row, 51:55])))
  n_xs3 <- mean(unlist(unname(numeric_combined_df[row, 56:60])))
  
  v_ms2 <- mean(unlist(unname(numeric_combined_df[row, 61:63])))
  v_xs2 <- mean(unlist(unname(numeric_combined_df[row, 64:66])))
  v_ms3 <- mean(unlist(unname(numeric_combined_df[row, 67:69])))
  v_xs3 <- mean(unlist(unname(numeric_combined_df[row, 70:72])))
  
  avg_row <- list(a_ms1, a_xs1, a_ms2, a_xs2, a_ms3, a_xs3, n_ms1, n_xs1, n_ms2, n_xs2, n_ms3, n_xs3, v_ms2, v_xs2, v_ms3, v_xs3)
  
  averaged_df <- rbind(averaged_df, avg_row)
}
colnames(averaged_df) <- c('a_m_s1', 'a_x_s1', 'a_m_s2', 'a_x_s2', 'a_m_s3', 'a_x_s3', 'n_m_s1', 'n_x_s1', 'n_m_s2', 'n_x_s2', 'n_m_s3', 'n_x_s3', 'v_m_s2', 'v_x_s2', 'v_m_s3', 'v_x_s3')

rownames(averaged_df) <- rownames(combined_df)

```

### Determine parameters for WGCNA

```{r}
tlog_averaged_df <- t(log2(averaged_df + 1))

sft <- pickSoftThreshold(tlog_averaged_df,
  dataIsExpr = TRUE,
  corFnc = cor,
  networkType = "signed"
)

sft_df <- data.frame(sft$fitIndices) %>%
  dplyr::mutate(model_fit = -sign(slope) * SFT.R.sq)

ggplot(sft_df, aes(x = Power, y = model_fit, label = Power)) +
  # Plot the points
  geom_point() +
  # We'll put the Power labels slightly above the data points
  geom_text(nudge_y = 0.1) +
  # We will plot what WGCNA recommends as an R^2 cutoff
  geom_hline(yintercept = 0.80, col = "red") +
  # Just in case our values are low, we want to make sure we can still see the 0.80 level
  ylim(c(min(sft_df$model_fit), 1.05)) +
  # We can add more sensible labels for our axis
  xlab("Soft Threshold (power)") +
  ylab("Scale Free Topology Model Fit, signed R^2") +
  ggtitle("Scale independence") +
  # This adds some nicer aesthetics to our plot
  theme_classic()
```

### Make WGCNA Dendrogram

```{r}
picked_power = 14
temp_cor <- cor       
cor <- WGCNA::cor         # Force it to use WGCNA cor function (fix a namespace conflict issue)
netwk <- blockwiseModules(tlog_averaged_df,                # <= input here

                          # == Adjacency Function ==
                          power = picked_power,                # <= power here
                          networkType = "signed",

                          # == Tree and Block Options ==
                          deepSplit = 2,
                          pamRespectsDendro = F,
                          # detectCutHeight = 0.75,
                          minModuleSize = 30,
                          maxBlockSize = 10000,

                          # == Module Adjustments ==
                          reassignThreshold = 0,
                          mergeCutHeight = 0.25,

                          # == TOM == Archive the run results in TOM file (saves time)
                          saveTOMs = T,
                          saveTOMFileBase = "ER",

                          # == Output Options
                          numericLabels = T,
                          verbose = 3)
cor <- temp_cor 

mergedColors = labels2colors(netwk$colors)
# Plot the dendrogram and the module colors underneath
plotDendroAndColors(
  netwk$dendrograms[[1]],
  mergedColors[netwk$blockGenes[[1]]],
  "Module colors",
  dendroLabels = FALSE,
  hang = 0.03,
  addGuide = TRUE,
  guideHang = 0.05 )

module_df <- data.frame(
  gene_id = names(netwk$colors),
  colors = labels2colors(netwk$colors)
)

#write_delim(module_df,
            #file = "gene_modules_nc_log_avg_wgcna.txt",
            #delim = "\t")
```

### Plot modules

```{r}
# Get Module Eigengenes per cluster
MEs0 <- moduleEigengenes(tlog_averaged_df, mergedColors)$eigengenes

# Reorder modules so similar modules are next to each other
MEs0 <- orderMEs(MEs0)
module_order = names(MEs0) %>% gsub("ME","", .)

# Add treatment names
MEs0$treatment = row.names(MEs0)

# tidy & plot data
mME = MEs0 %>%
  pivot_longer(-treatment) %>%
  mutate(
    name = gsub("ME", "", name),
    name = factor(name, levels = module_order)
  )


level_order <- c('a_m_s1', 'a_x_s1', 'a_m_s2', 'a_x_s2', 'a_m_s3', 'a_x_s3', 'n_m_s1', 'n_x_s1', 'n_m_s2', 'n_x_s2', 'n_m_s3', 'n_x_s3', 'v_m_s2', 'v_x_s2', 'v_m_s3', 'v_x_s3')
module_plot <- mME %>% ggplot(., aes(x=treatment, y=name, fill=value)) +
  geom_tile() +
  scale_x_discrete(limits = level_order) +
  theme_bw() +
  scale_fill_gradient2(
    low = "blue",
    high = "red",
    mid = "white",
    midpoint = 0,
    limit = c(-1,1)) +
  theme(axis.text.x = element_text(angle=90)) +
  labs(title = "Module-trait Relationships", y = "Modules", fill="corr")

module_plot
#ggsave("log_avg_module_heatmap.pdf", plot = module_plot, width = 8, height = 6)
```

### Make networks for each module

```{r}
library(igraph)

prrs <- read.delim("RGs.txt")


#grey = 1, red = 2, black = 3, magenta = 4, turquoise = 5, brown = 6, yellow = 7, green = 8, pink = 9, blue = 10
module <- "blue"  # Change to the name of your module
module_genes <- module_df$gene_id[module_df$colors == module]
prr_module_genes <- module_genes[module_genes %in% prrs$GeneID] #filter down to only PRRs

# Get the expression data for the module genes
exprDataModule <- tlog_averaged_df[, prr_module_genes]

# Compute the correlation matrix for the module genes
correlation_matrix_module <- cor(exprDataModule, method = "pearson")

# Define the threshold value for including edges
threshold <- 0.8  # Adjust as necessary

# Create the weighted adjacency matrix
# Retain correlation values above the threshold, set others to 0
adjacency_matrix_module <- ifelse(correlation_matrix_module > threshold, correlation_matrix_module, 0)

# Create a graph object using igraph with weights
network_module <- graph_from_adjacency_matrix(adjacency_matrix_module, mode = "undirected", weighted = TRUE)

edge_weights <- E(network_module)$weight
edge_weights_normalized <- (edge_weights - threshold) / (max(edge_weights) - threshold)  

# Define a color palette for the edges
color_palette <- colorRampPalette(c("lightblue", "darkblue"))  # Gradient from light blue to dark blue
edge_colors <- color_palette(100)[as.numeric(cut(edge_weights_normalized, breaks = 100))]  # Map normalized values to colors

# Remove self-loops
network_module <- delete_edges(network_module, which(ends(network_module, E(network_module))[,1] == ends(network_module, E(network_module))[,2]))




edge_attributes <- data.frame(
  source = ends(network_module, E(network_module))[,1],
  target = ends(network_module, E(network_module))[,2],
  weight = E(network_module)$weight
)

# Export edge attributes
write.csv(edge_attributes, file = "module10_edges.csv", row.names = FALSE)


node_attributes <- data.frame(
  id = V(network_module)$name
)

# Export node attributes to CSV
write.csv(node_attributes, file = "module10_nodes.csv", row.names = FALSE)




# Plot the network using a suitable layout
pdf("network_plot.pdf", width = 12, height = 10)
plot(network_module, layout = layout_with_kk(network_module), 
     vertex.label = V(network_module)$name, 
     vertex.size = 3, 
     edge.width = E(network_module)$weight * 1,  # Adjust scaling for better visibility
     edge.color = edge_colors)  # Optional: Adjust edge color and transparency
dev.off()


```

```{r}
correlation_matrix <- cor(tlog_averaged_df, method = "pearson")

# Convert correlation to adjacency matrix by applying a threshold
threshold <- 0.5  # Adjust to keep stronger connections
adjacency_matrix <- (correlation_matrix > threshold) * 1

network <- graph.adjacency(adjacency_matrix, mode = "undirected", weighted = TRUE)

# Plot the network using a suitable layout
plot(network, layout = layout_with_kk(network), vertex.label = V(network)$name, vertex.size = 5, edge.width = E(network)$weight * 10)
```
